---
title: "Brownlows Bets"
author: "Hailey Watkins-Oliver"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings
## the standard deviation and a 95% confidence interval for their earnings?

```{r}
mymin=numeric(1000)
for (i in 1:1000) {
    mymin[i]=BrLw()
}
smymin <- sort(mymin)

low=smymin[25]
high=smymin[975]

cat("Low value (25th percentile):", low, "\n")
cat("High value (975th percentile):", high, "\n")
```



# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero.

```{r}
mymin=numeric(1000)
for (i in 1:1000) {
  mymin[i]=BrLw(odds=4.389)
}
 smymin<- sort(mymin)
 
 low=smymin[25]
 high=smymin[975]
 
 mean(smymin)
 
  cat("low value (25th percentile):", low, "\n")
 cat("high value (975th percentile):", high, "\n")
 
 hist(smymin,
      main="Histogram of Sorted Values",
      xlab="Values",
      ylab="Frequency",
      col="skyblue",
      border="blue",
      breaks=100)
```

To determine how Brownlow's earnings reach zero, we identified an odds value of 4.389 through trial and error, as this value consistently yielded the most reliable results. To establish a 95% confidence interval, we calculated the lower and upper percentiles by marking the bottom and top 25 observations out of 1,000, leaving 950 observations (95% of the data) within the interval. This approach ensured a precise 95% confidence interval.

# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

```{r}
mymin=numeric(1000)
for (i in 1:1000) {
  mymin[i]=BrLw(numClass = 28)
}
 smymin<- sort(mymin)
 
 low=smymin[25]
 high=smymin[975]
 
 mean(smymin)
 
  cat("low value (25th percentile):", low, "\n")
 cat("high value (975th percentile):", high, "\n")
 
 hist(smymin,
      main="Histogram of Sorted Values",
      xlab="Values",
      ylab="Frequency",
      col="skyblue",
      border="blue",
      breaks=100)
```

Although 28 students is the ideal class size, Brownlow's revenues are still not quite nil. It would take a fraction of pupils to reach negative incomes, which is not possible. Approximately $29 is earned at a class size of 29, and -$23 is earned at a class size of 28. Therefore, the optimum way to get Brownlow's earnings as near to zero as feasible is to have a class size of 28. It was trial and error that led to this conclusion.

# Discussion and Conclusion

We were able to demonstrate how Brownlow loses money from students because of the 2:1 odds using BrLw.R. We could alter the chances and the class size to make the earnings zero in order to make this zero. Has 28 students in the class and chances of 4.389. 


